A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations

### 2019-02-11 paper.tex
Updates from Overleaf

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory \todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away. We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

One-second-cadence geoelectric field and geomagnetic field data from Memambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}. Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.

@@

The GIC data contains non-physical spikes that are followed by a relaxation time of  40-80 seconds. These spikes were removed by identifying times $t_s$ when the change in 1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ were replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.

\section{Models}

To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model~1,

@@

\noindent
the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$. \citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}

In Model 2, the system parameters are frequency dependent

@@

Note that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantity $\mathbf{B}$ or its components, $B_x$ and $B_y$ refer to the magnetic field.

The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model2}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:

@@

\noindent
where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives

@@

That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field. In this work, we have computed $\mathbf{Z}$ using the method described later in this section.

A fourth model considered is:

@@

\noindent
In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produce predictions of $GIC$ that are similar to that obtained from Model 3.

For frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$. To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed. The transfer function magnitudes $|A(\omega)|$ and $|B(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude). The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$. Additional details on the method are given in \cite{Weigel2017} and references therein.

@@

Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data. The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components of the 35 transfer functions. The average impulse response functions were obtained by averaging the 35 impulse response functions.

There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic least-squares method \citep{Simpson2005}. These works typically involve the estimation of a robust and bias-corrected transfer function. The results presented used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters. As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

However, several alternative methods for computing the transfer functions were tested individually to determine if there was a change in the computed transfer functions and, most important for this work, each model's out-of-sample prediction ability. When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance are within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$3 (corresponding to a period of $\sim 30$ seconds. The experiments in this regard are discussed in the following section and in the Appendix.

@@

The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.

@@

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\end{table}

@@

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1. Model 3, which uses magnetic field measurements as in input provides better estimates of GIC than Model 2. Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.

Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models. The error bars extend by $\pm$ one standard error of the 35 SN ratios at each period. The SN ratio for Model 1 is always lower than that for Models 2-4.  Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

@@

The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of GIC \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GIC, but is a poor approximation relative to Models 2 and 3. For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition). Note that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2. In Model~2, GIC is most dependent on $E_y$, however note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}) so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.

@@

Although the primary objective of this work is the creation of the models for comparison of the predictive ability, a natural outcome is model parameter estimates. As noted in the introduction, the motivation for using a basic least squares method is (1) that it reduces the number of adjustable methodological parameters that must be considered and (2) because we are attempting to consider which model performs best in a least-square-error sense. It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used. Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable in the sense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).

@@

Because Model~4 provided the best prediction performance, the ideal measurements to use for estimating GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. GIC data provided by ... \todo[color=blue!40]{Shinichi: How should this be referenced?}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_H-options-1.pdf}
\caption{Impulse responses for Models 1, 2, and 4. The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4. The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
\label{H}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_Z-options-1.pdf}
\caption{Frequency domain transfer functions for Models 1, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
\label{Z}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
\caption{Frequency domain phase values for Models 1, 2, and 4.}
\label{Phi}
\end{figure}

@@

\item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day. If $u(D)$ and $v(D)$ are the real and imaginary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation). The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$. Neither averaging method was yielded consistently higher or lower values.

>>>>

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100 km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10\todo[color=blue!40]{Shinichi: Is 10 km correct?} $\sim${\color{red}\textbf{10}} kilometers away.  We compute frequency dependent network parameters  $A(\omega)$ and $B(\omega)$ in Model 2, $GIC(\omega)$ $=$ $A(\omega)E_x(\omega) + B(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

One-second-cadence geoelectric field and geomagnetic field data from KakiokaMemambetsu Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.

Unphysical s Data from 2001-2011 have been used by ~\cite{Fujii2015} to estimate the magnetotelluric impedance at the site.

@@

The GIC data contains non-physical spikes that are followed by a relaxation time of approximately  40-80 seconds. These spikes were removed by identifying times $t_s$ when the change in $|GIC|$1-second $|GIC|$ measurements was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ wasere replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.

\section{Models and Methods}

To determine the coefficients $a_o$ and $b_o$ in the frequency independent model, Model ~1,

@@

\noindent
we form the matrix equation $GIC = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $GIC$ is a 864000x1 matrix of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with $N$ rows of $[E_x(t), E_y(t)]$ observations. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ for all values of $t$. \citep[][ provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}

In Model 2, we assume that the system parameters are frequency dependent

@@

\noindent
In the time domain, this is equivalent to

\begin{linenomath*}
\begin{equation}
G_E(t) = \sum_{\tau=-\infty}^{\infty}\big[a(\tau)E_x(t-\tau) + b(\tau)E_y(t-\tau)\big]
\label{model2b}
\tag{2b}
\end{eNote that the scalar quantity $B$ without a subscript refers to the system parameter in Model~2 while the vector quantion}
\end{linenomath*}

\noindent
were the parameters $a$ and $A$ and $bty $\mathbf{B}$ or its components, $B_x$ and $B_y$ are inverse fourier transforms of each otherfer to the magnetic field.

The third model considered uses the electric field $\mathbf{E}'$ in Equation~\ref{model12}, where $\mathbf{E}'$ was computed using the measured magnetic field and a transfer function $\mathbf{Z}$: $\mathbf{E}'(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$:

@@

\noindent
where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xyx}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. Substitution gives

@@

That is, instead of using the measured electric field directly as in Model 2, we use a transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ to provide an estimate of the electric field based on the magnetic field. In this work, we have computed $\mathbf{Z}$ using the method described later in this section; $A(\omega)$ and $B(\omega)$ are taken from the solution of Equation~\ref{model2}, using the method also described later in this section.

A fourth model was also considered is:

@@

\noindent
In this model, the transfer function components $Z_x$ and $Z_y$ are determined directly from measurements of $GIC$ and $\mathbf{B}$ and the model is expected to produced predictions of $GIC$ that are similar to that obtained from Model 3.

The parameters in Model 2 and 4 can be computed using either a time or frequency domain method. In the time domain method, fixed values of the upper ($N_c$) and lower ($N_c$) limits are selected in the summations in Equation~\ref{model2b} (and equivalent for Equation 4) and then the equation is repeated for at least ($N_c+N_a+1$) time values and then the resulting over-determined set of equations is inverted to find the unknowns $a(\tau)$ and $b(\tau)$ for $N_a \le \tau \le Nc$.

In theFor frequency domain method, estimates of $A(\omega)$ and $B(\omega)$ are made at evaluation frequencies, $\omega_e$, by performing a least squares regression using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $\omega_e$. To convert the frequency domain transfer functions $A(\omega)$ and $B(\omega)$ estimated at the frequencies $\omega_e$ to impulse responses $a(\tau)$ and $b(\tau)$, they are first linearly interpolated onto a 1 Hz frequency grid and then inverse fourier transformed. The transfer function magnitudes $|A(\omega)|$ and $|B(\omega)|$ were calculated by linear interpolation of the transfer function magnitudes (instead of linearly interpolating the real and imaginary components and then computing the magnitude). The same approach is used for solving for the transfer function $\mathbf{Z}$ in $\mathbf{E}(\omega) = \mathbf{Z}(\omega)\mathbf{B}(\omega)$ - the $xx$ and $yx$ components of $\mathbf{Z}$ are solved for using measured values of $E_x$, $B_x$, and $B_y$ in $E_x = Z_{xx}B_x + Z_{yx}B_{y}$, and the $yx$ and $yy$ components of $\mathbf{Z}$ are solved for using measured values of $E_y$, $B_x$, and $B_y$ in $E_y = Z_{yx}B_x + Z_{yy}B_y$. Additional details on the method are given in \cite[][{Weigel2017} and references therein.]{Weigel2017}.

@@

Transfer function parameters were computed for each of the 35 unique 1-day window segments of available data. The average transfer function magnitude was obtained by averaging these 35 transfer function magnitudes; the average transfer function phase was obtained using the averaged real and imaginary components, respectively, of the 35 transfer functions. The average impulse response functions were obtained by averaging the 35 impulse response functions.

There are many statistical methods that have been proposed to estimate transfer functions in the magnetotelluric literature that use additional steps, processing, and assumptions \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Simpson2005}. The additional steps and processing were developed to address leakage, un-removed data spikes and shifts, high levels of noise contamination, time intervals in which the model assumptions are violated, bias, and the fact that errors in the fits of the transfer function at each evaluation frequency may not be Gaussian distributedleast-squares method \citep{Simpson2005}. These works typically involve the estimation of a robust and low bias-corrected transfer function. In tThe results presented, we did not use a robust statistical method to calculate the transfer function or a bias correction algorithm - used were obtained using the basic least-squares transfer function estimation method as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating low-bias model parameters. As discussed in ~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

However, several alternative methods for computing the transfer functions were tested individually by modifying the basic method described above to determine if there was a change in the computed transfer functions and each model's, most important for this work, each model's out-of-sample prediction ability. When using these alternative methods, we find that the conclusions made regarding the differences each model's prediction ability do not change, the differences in each model's prediction performance isare within the calculated uncertainty of the used method, and the transfer function values are generally within the error bars of the used method when the signal to noise ratio is above $\sim$53 (corresponding to a period of $\sim 30$ seconds. The experiments in this regard are discussed in the following section and in the Appendix.

@@

The performance of each model in predicting $GIC$ was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.

@@

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. $MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\end{table}

@@

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1. Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2. Model 4, in which a transfer function that connects $GIC$ to $\mathbf{B}$ was derived directly from the data, provides the best predictions of $GIC$.

Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models. The error bars representextend by $\pm$ one standard error of the 35 SN ratios at each period. Consistent with the results in Table~1, tThe SN ratio for Model 1 is consistentlyalways lower than that for Models 2-4.  Above $1400$~s (24~min), the SN ratio ordering is consistent with the ordering of the PE and MSE ratios shown in Table~1. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

@@

The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$  \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3. For the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

@@

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition). For this systemNote that the ratio of $|\overline{b}_o/||\overline{a}_o|$ is approximately 1 for Model~1 but varies from 1.6 to 6.3 for Model~2. In Model~2, GIC is most dependent on $E_y$, whichhowever note that $E_y$ has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).

Figure~\ref{Phi} shows the transfer function phases. By definition, that for Model 1 is zero.

Several of these methods were tested individually by modifying the method described above to determine their impact on the computed transfer functions and their over all results.

\begin{itemize}
\item Using robust regression (with either bi-square or Huber weighting) instead of linear regression when comput so that $E_x$ and $E_y$ on average contribute approximately equal amounts to the predictability of GIC.

@@

Although the primary objective of this work is the creation of the models for comparison of the predictive ability, a natural outcome is model parameter estimates. As noted in the introduction, the motivation for using a basic least squares method is (1) that it reduces the number of adjustable methodological parameters that must be considered and (2) because we are attempting to consider which model performs best in a least-square-error sense. It is natural to ask if the model parameter estimates are biased and how sensitive they depend on the statistical method used. Several alternative estimation methods have been considered and are discussed in the Appendix; the primary conclusion is that the values shown in Table~1 are not sensitive and the parameter estimates are only sensitive below approximately $T=30$~s (they are unreliable ing the transfer function estimates at each $\omega_e$

\item Averaging the 35 transfer functions using the Huber M-estimator with a $\psi$-function \citep{Huber2011}.

\item Using weighted averages of the 35 transfer functions with the weighting at each frequency equal to the in-sample coherence at that frequency

\item Windowing the signals in the frequency dosense that the alternative estimation methods give parameter values outside of the error bars shown for the baseline method).

@@

Because Model~4 provided the best prediction performance, the ideal measurements to use for estimatin using a rectangular window (instead of Parzen)
\end{itemize}

\section{Discussion and Conclusions}
g GIC in this dataset are contemporaneous GIC and $\mathbf{B}$ measurements.

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. GIC data provided by ...

\bibliography{paper.bib}
 \todo[color=blue!40]{Shinichi: How should this be referenced?}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_SN-options-1.pdf}
\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_H-options-1.pdf}
\caption{Impulse responses for Models 1, 2, and 4. The circles at $t = 0$ correspond to the impulse response for Model 1; the red/blue lines are the impulse responses for Model 2 and 4. The values of $a_o$ and $b_o$ have been scaled by a factor of 10 for readability.}
\label{H}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summaryTF_aves_Z-options-1.pdf}
\caption{Frequency domain transfer functions for Models 1-3, 2, and 4. By definition, the frequency domain transfer functions for $A_o$ and $B_o$ are constant and equal to $a_o$ and $b_o$.}
\label{Z}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_summary_Phi-options-1.pdf}
\caption{Frequency doTF_aves_Phi-options-1.pdf}
\caption{Frequency domain phase values for Models 1, 2, and 4.}
\label{Phi}
\end{figure}

@@

\item Five methods for averaging were used to compute the magnitudes at each evaluation frequency of model parameter $p(D)$, where $D$ is estimate made using a 1-day. If $u(D)$ and $v(D)$ are the real and imagin phase values for Models 1-3.}
\label{Phi}
\end{figure}
ary components of $p$, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation). The PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$. Neither averaging method was yielded consistently higher or lower values.
